### 2025-08-01 review/response_250623.tex

					\caption{Theoretical singular value outer bounds \eqref{iq:horn} (uniformly-spaced mesh grids) vs achievable singular value region by solving \eqref{op:pareto} (solid dark shape) for one channel realization, where $N_\mathrm{T}=N_\mathrm{S}=N_\mathrm{R}=3$, the direct channel is negligible, and the \gls{bd}-\gls{ris} is fully-connected. Small offsets are introduced on both views such that the active bounds are highlighted by densely-spaced curves/lines that marginally overlap the region from above. The achievable region lies entirely within the intersection of the bounding surfaces in the 3D space.}
					\label{fg:singular_region}
				\end{figure}

				Fig. \ref{fg:singular_region} compares the achievable singular value region obtained by solving problem \eqref{op:pareto} and its outer bounds suggested by Corollary \ref{co:nd_sv_prod_subset}.
				Here $\bar{N}=N_\mathrm{T}=N_\mathrm{S}=N_\mathrm{R}=3$ and the bounds are enumerated as
				\begin{subequations}
					\begin{equation}
						\label{iq:sv_3_individual}
						\begin{gathered}
							\sigma_1(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}), \quad
							\sigma_2(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_2(\mathbf{H}_\mathrm{F}), \\
							\sigma_2(\mathbf{H}) \le \sigma_2(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}), \quad
							\sigma_3(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_3(\mathbf{H}_\mathrm{F}), \\
							\sigma_3(\mathbf{H}) \le \sigma_2(\mathbf{H}_\mathrm{B}) \sigma_2(\mathbf{H}_\mathrm{F}), \quad
							\sigma_3(\mathbf{H}) \le \sigma_3(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}),
						\end{gathered}
					\end{equation}
					\begin{equation}
						\label{iq:sv_3_product}
						\begin{gathered}
							\sigma_1(\mathbf{H}) \sigma_2(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_2(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}) \sigma_2(\mathbf{H}_\mathrm{F}), \\
							\sigma_1(\mathbf{H}) \sigma_3(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_3(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}) \sigma_2(\mathbf{H}_\mathrm{F}), \\
							\sigma_1(\mathbf{H}) \sigma_3(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_2(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}) \sigma_3(\mathbf{H}_\mathrm{F}), \\
							\sigma_2(\mathbf{H}) \sigma_3(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_2(\mathbf{H}_\mathrm{B}) \sigma_2(\mathbf{H}_\mathrm{F}) \sigma_3(\mathbf{H}_\mathrm{F}), \\
							\sigma_2(\mathbf{H}) \sigma_3(\mathbf{H}) \le \sigma_2(\mathbf{H}_\mathrm{B}) \sigma_3(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}) \sigma_2(\mathbf{H}_\mathrm{F}), \\
							\sigma_2(\mathbf{H}) \sigma_3(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_3(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}) \sigma_3(\mathbf{H}_\mathrm{F}),
						\end{gathered}
					\end{equation}
					\begin{equation}
						\label{iq:sv_3_ordering}
						\sigma_1(\mathbf{H}) \ge \sigma_2(\mathbf{H}) \ge \sigma_3(\mathbf{H}),
					\end{equation}
				\end{subequations}
				where \eqref{iq:sv_3_individual}, \eqref{iq:sv_3_product} are explicit results of \eqref{iq:horn} while \eqref{iq:sv_3_ordering} denotes the ordering of singular values.
				Those are labeled respectively as `Bounds (individual)', `Bounds (product)', and `Bounds (ordering)' in Fig. \ref{fg:singular_region}.
				The two views confirm that the theoretical outer bounds are not everywhere tight with many entries being redundant, but they provide a conservative estimate of the achievable singular value region.
				Importantly, the vertices of the region are always on the boundary, which implies that those can be achieved in the closed form without performing optimization.
			}
			\item The bounds in Corollary \ref{co:nd_sv_prod_tail} reveal the upper (resp. lower) bound on the product of some largest (resp. smallest) channel \glspl{sv}. These bounds can be applied, for instance, as a shortcut to establish the upper bound of \gls{bd}-\gls{ris}-aided \gls{mimo} channel capacity at extreme \glspl{snr}.
			\change{
				Corollary \ref{co:nd_sv_prod_tail} reveals the shaping limits on the product of some extreme channel singular values.
				The lower bounds \eqref{iq:sv_nd_prod_smallest} coincide at zero when $\bar{N} \ne N$ (i.e., $N_\mathrm{T} = N_\mathrm{S} = N_\mathrm{R}$ being false).
				These bounds can be applied, for instance, as a shortcut to establish the capacity of \gls{bd}-\gls{ris}-aided \gls{mimo} channels at extreme \gls{snr}, as shown in Corollary \ref{co:nd_capacity_snr_extreme}.
				In the special case $k=1$, we arrive at the upper bound on the largest channel singular value $\sigma_1(\mathbf{H}) {\le} \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F})$.
				This is particularly useful for \gls{mimo} wireless power transfer with \gls{rf} combining where the harvested power depends merely on, and is a quartic function of, the largest channel singular value \cite{Shen2021}.
				A closed-form \gls{bd}-\gls{ris} solution to attain this upper bound can be found below.
			}
			\item The bounds in Corollary \ref{co:nd_sv_indl} reveal the shaping limits of the $n$-th largest channel \gls{sv}. These bounds enable closed-form passive beamforming, and hence fixed channel and closed-form active beamforming, for spatial multiplexing with a limited number $n$ of \gls{rf} chains. Another use case is to enhance the harvested power for \gls{mimo} wireless power transfer with \gls{rf} combining by maximizing the dominant channel \gls{sv}.
			One may also improve the channel condition number for numerically stable signal processing by maximizing the smallest channel \gls{sv}.
			\change{
				Corollary \ref{co:nd_sv_indl} and Proposition \ref{pp:rd} both reveal the shaping limits of the $n$-th largest channel singular value.
				% , which may be used to simplify the \gls{mimo} precoder design with limited number $n$ of \gls{rf} chains.
				They are derived under different assumptions are not special cases of each other.
				Importantly, Corollary \ref{co:nd_sv_indl} establishes upper and lower bounds for \emph{each} channel singular value (c.f. first and last $k$ in Proposition \ref{pp:rd}) and provides general solutions for fully-connected \gls{bd}-\gls{ris} of arbitrary (c.f. sufficiently large) size to attain the equalities.
				% These bounds can be applied, for instance, to provide a closed-form passive beamforming (and thus fixed channel and closed-form active beamforming) for spatial multiplexing with a limited number $n$ of \gls{rf} chains.
				These bounds enable closed-form passive beamforming, and hence fixed channel and closed-form active beamforming, for spatial multiplexing with a limited number $n$ of \gls{rf} chains.
				% Another use case is to enhance the harvested power for \gls{mimo} wireless power transfer with \gls{rf} combining by maximizing the dominant channel singular value \cite{Shen2021}.
				% One may also improve the channel condition number for numerically stable signal processing by maximizing the smallest channel singular value.
				We emphasize that in \eqref{eq:ris_nd_sv_indl} the mode alignment is realized by $\mathbf{V}_\mathrm{B}$ and $\mathbf{U}_\mathrm{F}$ while the ordering is enabled by permutation matrices $\mathbf{P}$ and $\mathbf{Q}$, which are special cases of $\mathbf{X}$ defined in \eqref{eq:ris_decompose_group}.
			}
			\item While \cite{Rong2009a} discussed the optimal beamforming structure for multi-hop \gls{af} relays, its focuses of channel diagonalization and relay power allocation are fundamentally different from this paper. The literature provides neither channel \gls{sv} bounds nor channel capacity bounds. Please refer to our response to the editor on this point.
		\end{itemize}
	}

@@

	\comment{
	Prop. \ref{pp:shaping} imposes the assumption that $f(\cdot)$ must be symmetric gauge function in \glspl{sv} of $\mathbf{H}$. Note gauge function is strong assumption since it should be homogeneous and convex (e.g., \cite[Sec 3-I]{Marshall2010}) in \glspl{sv} of $\mathbf{H}$, which indeed severely restricts the range of its application. When $f(\cdot)$ is nonconvex, subdifferential ($\mathbf{D}$ in eq. \eqref{eq:shaping_subdiff}) generally does not exist. In fact, any commonly adopted performance metrics, e.g., spectral efficiency, energy efficiency, receiving power are not gauge functions (they are inhomogeneous and nonconvex).
	On the other hand, to obtain partial differential with respect to (w.r.t.) \gls{ris}, (i.e., the aim of Prop. \ref{pp:shaping}), the assumption of gauge function is unnecessary. Directly taking the derivative of $f(\cdot)$ w.r.t. \gls{ris} is more preferable.
	}

	\reply{
		We fully agree with the reviewer that enforcing $f(\mathrm{sv}(\mathbf{H}))$ as a gauge function is too restrictive and unnecessary for the shaping problem.
		In the revised manuscript, Prop \ref{pp:shaping} has been extended to locally Lipschitz functions of channel singular values.
		We sincerely appreciate your insights on this issue.

		\change{
			Below we propose a numerical method to optimize \gls{bd}-\gls{ris} for a broad class of singular value functions.
			\setcounter{definition}{2}
			\begin{definition}[Locally Lipschitz]
				A function $f: \mathbb{R}^N \to \mathbb{R}$ is locally Lipschitz if for any compact set $\mathcal{S} \subset \mathbb{R}^N$, there exists a constant $L > 0$ such that $\lvert f(\mathbf{x}) - f(\mathbf{y}) \rvert \le L \lVert \mathbf{x} - \mathbf{y} \rVert$ for all $\mathbf{x}, \mathbf{y} \in \mathcal{S}$.
			\end{definition}

			\setcounter{proposition}{3}
			\begin{proposition}
				\label{pp:shaping}
				Consider channel shaping problems of the form
				\begin{maxi!}
					{\scriptstyle{\mathbf{\Theta}}}{f\bigl(\sv(\mathbf{H})\bigr)}{\label{op:shaping}}{\label{ob:shaping}}
					\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{cn:shaping_unitary}}
				\end{maxi!}
				where $f: \mathbb{R}^{N} \to \mathbb{R}$ is arbitrary locally Lipschitz function of channel singular values.
				The Clarke subdifferential of \eqref{ob:shaping} with respect to \gls{bd}-\gls{ris} block $g$ is
				\begin{equation}
					\label{eq:shaping_subdiff}
					\partial_{\mathbf{\Theta}_g^*} f\bigl(\sv(\mathbf{H})\bigr) = \conv \bigl\{ \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathbf{D} \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H} \bigr\},
					% \partial_{\mathbf{\Theta}_g^*} f\bigl(\sv(\mathbf{H})\bigr) = \conv \bigl\{ \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathbf{D} \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H} \bigr\},
				\end{equation}
				where $\mathbf{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$ is a rectangular diagonal matrix with $[\mathbf{D}]_{n,n} \in \partial_{\sigma_n(\mathbf{H})} f\bigl(\sv(\mathbf{H})\bigr)$, $\forall n \in [N]$, and $\mathbf{U}$, $\mathbf{V}$ are any left and right singular matrices of $\mathbf{H}$.
			\end{proposition}

			\begin{proof}
				Please refer to Appendix~\ref{ap:shaping}.
			\end{proof}

			Proposition \ref{pp:shaping} enables subgradient-based optimization for arbitrary locally Lipschitz function of channel singular values (e.g., Pareto frontier, power gain, capacity, condition number) via \eqref{eq:shaping_subdiff}.
			Next, we introduce a \emph{geodesic \gls{rcg}} method modified from \cite{Abrudan2008,Abrudan2009} for the optimization of \gls{bd}-\gls{ris}.
			Our contribution is an extension to the block-unitary case with sequential, parallel, or unified updates for accelerated convergence.
			The steps for updating $\mathbf{\Theta}_g$ at iteration $r$ are summarized below, where the gradients are replaced by Clarke subgradients for non-smooth $f$.
		}
	}

@@

	\comment{
	The authors propose Riemannian Conjugate Gradient (\gls{rcg}) algorithm (Algorithm \ref{ag:rcg}) to optimize \gls{ris} configuration. Note CG descent on manifold is indeed well established and is widely used for optimization on manifolds, e.g., \cite{Boumal23t}. It can be implemented conveniently by off-the-shelf solvers, e.g., Manopt \cite{Boumal14a}. Then, what is the novelty or significance of Sec. \ref{sc:shaping_numerical}?
	}

	\reply{
		Manopt is indeed a canonical tool for manifold optimization. Upon inspecting the source code, we found it invokes a generic non-geodesic solver for feasible domain \texttt{stiefelcomplexfactory(n, p, k)} that does not exploit the block-unitary structure of our problem. This is also confirmed by MATLAB profiler that $~60\%$ of its running time was spent on projection (i.e. retraction from the Euclidean space to the manifold, due to non-geodesic update). This is the main bottleneck of the algorithm.
		\begin{figure}[H]
			\centering
			\includegraphics[width=0.75\columnwidth]{../assets/review/profiler.png}
			\caption{MATLAB profiler report of Manopt vs proposed \gls{rcg} on 1000 independent channel realizations. The simulation parameters are inline with Table \ref{tb:complexity_algorithm}.}
		\end{figure}

		In contrast, our proposed geodesic \gls{rcg} algorithm is tailored to group-connected \gls{bd}-\gls{ris} designs, offering significant computational advantages as discussed in context. The novelty of Section \ref{sc:shaping_numerical} lies in:
		\begin{itemize}
			\item The formulation of the channel shaping problem as an optimization problem on the product of unitary manifolds, which is a natural and appropriate setting for \gls{bd}-\gls{ris} design.
			\item The derivation of the translated Riemannian gradient at the identity of the manifold, which offers a compact expression for the geodesic update.
			\item The development of a specialized geodesic \gls{rcg} algorithm that leverages the block-unitary structure of \gls{bd}-\gls{ris} for accelerated convergence and reduced computational complexity compared to general-purpose manifold optimization solvers.
			\item The application of this algorithm to reveal the entire achievable channel \gls{sv} region, which is a novel contribution to the understanding of \gls{bd}-\gls{ris} shaping capabilities.
		\end{itemize}

		In the revised manuscript, we have added Table \ref{tb:complexity_algorithm} to benchmark the performance of two \gls{rcg} algorithms on the Pareto channel \gls{sv} problem \eqref{op:pareto}

		\change{
			\begin{table}[H]
				\caption{Performance of \gls{rcg} Algorithms on \eqref{op:pareto} with $N_\mathrm{T}=N_\mathrm{R}=4$, $L=16$}
				\label{tb:complexity_algorithm}
				\centering
				\resizebox{\columnwidth}{!}{
					\begin{tabular}{ccccccc}
						\toprule
						\multirow{2}{*}{\gls{rcg} path} & \multicolumn{2}{c}{$N_\mathrm{S}=16$} & \multicolumn{2}{c}{$N_\mathrm{S}=64$} & \multicolumn{2}{c}{$N_\mathrm{S}=256$}                                      \\
						\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
						                                & Iterations                            & Time [ms]                             & Iterations                             & Time [ms] & Iterations & Time [ms] \\
						\midrule
						Geodesic                        & 6.493                                 & 1.807                                 & 9.003                                  & 7.378     & 12.98      & 49.41     \\
						Non-geodesic (Manopt)           & 8.601                                 & 25.90                                 & 11.09                                  & 36.27     & 14.29      & 65.89     \\
						\bottomrule
					\end{tabular}
				}
			\end{table}

### 2025-08-01 review/response_250623.tex

			Table \ref{tb:complexity_algorithm} benchmarks two \gls{rcg} algorithms on the Pareto singular value problem \eqref{op:pareto} with $N_\mathrm{T}=N_\mathrm{R}=4$ and $L=16$.
			The geodesic \gls{rcg} is implemented with pinched gradients w.r.t. $\mathbf{\Theta}$ and unified step size selection; please refer to the discussion below \eqref{eq:geodesic_tran} for details.
			The non-geodesic \gls{rcg} is implemented by \texttt{Manopt} toolbox at commit \texttt{a879a0d} \cite{Boumal14a}.
			Both algorithms employ a stopping criterion of relative change in the objective function with a tolerance of $\epsilon = \num{1e-4}$, such that the final values are identical within reasonable precision.
			The statistics are averaged over 1000 independent channel realizations.
			We observe that the non-geodesic \gls{rcg} typically requires 1 to 2 more iterations than its geodesic counterpart.
			This is because the addition is in the tangent space of the manifold and is less effective than manifold-native updates.
			When it comes to elapsed time, the geodesic \gls{rcg} is $1333\%$ faster than the non-geodesic counterpart when $N_\mathrm{S}=16$.
			The main reason is that the geodesic \gls{rcg} avoids the retraction step from the Euclidean space to the manifold.
			According to the profiler report, around $60\%$ of the non-geodesic \gls{rcg} runtime is spent on retraction, which becomes the main bottleneck of the algorithm.
			The advantage narrows down to $391.6\%$ and $33.35\%$ when $N_\mathrm{S}=64$ and $256$, respectively.
			This is because accurately evaluating matrix exponential can be time-consuming for large $N_\mathrm{S}$.
		}
	}
